---
title: "EDA - Global"
---

# Setup
## Packages
```{r, message = FALSE}
library(tidyverse)
library(tidycovid19)
library(zoo)
library(knitr)
library(kableExtra)
library(naniar)
```

## Data
Let's do some feature engineering and look at the data structure:
```{r}
df <- download_merged_data(cached = TRUE, silent = TRUE) %>%
  group_by(country) %>%
  mutate(
    # New daily confirmed cases
    new_confirmed = confirmed - lag(confirmed),
    # New daily confirmed cases per capita
    new_confirmed_pc = new_confirmed / population,
    # Cum confirmed cases per capita
    confirmed_pc = confirmed / population,
    
    # New daily deaths
    new_deaths = deaths - lag(deaths),
    # New daily deaths per capita
    new_deaths_pc = new_deaths / population,
    # Cum deaths per capita
    deaths_pc = deaths / population,
    
    # 7 Day moving averages for confirmed, deaths, and per capita
    ave_7d_new_confirmed = rollmean(new_confirmed, 7, na.pad=TRUE, align="right"),
    ave_7d_new_deaths = rollmean(new_deaths, 7, na.pad=TRUE, align="right"),
    ave_7d_new_confirmed_pc = rollmean(new_confirmed_pc, 7, na.pad=TRUE, align="right"),
    ave_7d_new_deaths_pc = rollmean(new_deaths_pc, 7, na.pad=TRUE, align="right")
  ) %>%
  ungroup()

glimpse(df, width = 100)

```
## Variable Definitions
The package provides some definitions for the variables:
```{r}
var_definitions <- tidycovid19_variable_definitions %>%
  select(var_name, var_def)

kable(var_definitions) %>% 
  kableExtra::kable_styling() %>%
  kableExtra::scroll_box(width = "100%", height = "300px")
```
\

# Missing values analysis
```{r}
vis_miss(df[,1:10])
vis_miss(df[,11:20])
vis_miss(df[,21:30])
vis_miss(df[,31:40])
vis_miss(df[,41:50])

```
\
We can see that some variables have a substantial amount of values missing. When comparing countries, we should check the specific availability of the variable for the specific country.
  
The package hast variables that show numbers of reported measures. Some variables of interest that have only 11.47% missing data are the following:

* `soc_dist`
* `mov_rest`
* `pub_health`
* `gov_soc_econ`
* `lockdown`

Also of interest would be a time series of `total_tests`, `total_vaccinations`, `hosp_patients`, and `icu_patients`.
Due to the large number of NA's, an interpretation might not be very meaningful.
  
Further analysis per country might provide better data.

# EDA - Exploratory Data Analysis

## Global Numbers
Here a new `global_df` is generated for more control over the rolling mean and missing values.

### Data
```{r}
global_df <- df %>%
  group_by(date) %>%
  summarize(
    global_confirmed = sum(confirmed, na.rm = TRUE), 
    global_deaths = sum(deaths, na.rm = TRUE),
    global_soc_dist = sum(soc_dist, na.rm = TRUE),
    global_mov_rest = sum(mov_rest, na.rm = TRUE),
    global_lockdown = sum(lockdown, na.rm = TRUE)
    ) %>%
  mutate(
    global_new_confirmed = global_confirmed - lag(global_confirmed),
    global_new_deaths = global_deaths - lag(global_deaths),
    global_ave_7d_new_deaths = rollmean(global_new_deaths, 7, na.pad=TRUE, align="right"),
    global_ave_7d_new_confirmed = rollmean(global_new_confirmed, 7, na.pad=TRUE, align="right")
  )

glimpse(global_df)
```

### Global Confirmed Cases
```{r}
ggplot(global_df, aes(x = date)) +
  geom_col(aes(y = global_confirmed), alpha = 0.3) +
  labs(title = "Global total confirmed cases", x = "", y = "")
  
ggplot(global_df, aes(x = date)) +
  geom_col(aes(y = global_new_confirmed), alpha = 0.3) +
  geom_line(aes(y = global_ave_7d_new_confirmed), color = "red") +
  labs(title = "7-day rolling mean\nGlobal daily cases", x = "", y = "") +
  ylim(0,800000)
```

### Global Deaths
```{r}
ggplot(global_df, aes(x = date)) +
  geom_col(aes(y = global_deaths), alpha = 0.3) +
  labs(title = "Global total deaths", x = "", y = "")

ggplot(global_df, aes(x = date)) +
  geom_col(aes(y = global_new_deaths), alpha = 0.3) +
  geom_line(aes(y = global_ave_7d_new_deaths), color = "red") +
  labs(title = "7-day rolling mean\nGlobal daily deaths", x = "", y = "")
```
\

### Other Global Variables
```{r}
# Number of tests, vaccinations, and patients
ggplot(df, aes(x = date, y = total_tests), group = date) +
  geom_col() +
  labs(title = "Global total tests")

ggplot(df, aes(x = date, y = total_vaccinations), group = date) +
  geom_col() +
  labs(title = "Global total vaccinations")

ggplot(df, aes(x = date, y = hosp_patients), group = date) +
  geom_col() +
  labs(title = "Global total hospital patients")


ggplot(df, aes(x = date, y = icu_patients), group = date) +
  geom_col() +
  labs(title = "Global total icu patients")

# Number of measures reported
ggplot(global_df, aes(x = date, y = global_soc_dist)) +
  geom_col() +
  labs(title = "Global total social distancing measures")

ggplot(df, aes(x = date, y = mov_rest), group = date) +
  geom_col() +
  labs(title = "Global total movement restriction measures")

ggplot(df, aes(x = date, y = pub_health), group = date) +
  geom_col() +
  labs(title = "Global total public health measures")

ggplot(global_df, aes(x = date, y = global_lockdown), group = date) +
  geom_col() +
  labs(title = "Global total lockdown measures")

```

